Application of Network Flow and Linear Programming to Scheduling of Clinicians

### 2019-08-26 paper/paper.intro.tex
Work on intro and discussion sections
- Rewrite 3 goals in intro
- Add outline of paper to intro
- Address goals in separate paragraphs in discussion

%When generating these schedules, there are many variables and constraints that need to be taken into account, such as preventing multiple consecutive assignments for a given clinician or ensuring that assignments are spread out evenly throughout a certain period of time. It is very common for departments to opt for a manual method of generating these schedules [??]. Unfortunately, such methods are unreliable and can lead to unfavourable assignments, since a manual method may miss certain constraints or forget to account for a certain variable during the process. \\

Automated methods to optimize schedules have been studied and applied in many industries, including transportation [??], manufacturing [??], [...]. Of special interest to a clinician on-call scheduling problem are the approaches to schedule nurses, who often work in shifts. In the nurse scheduling problem, the goal is to find an optimal assignment of nurses to shifts that satisfies all of the hard constraints, such as hospital regulations, and as many soft constraints as possible, which may include nurse preferences. A wide variety of approaches, including exact and heuristic approaches, have been used to solve the nurse scheduling problem: integer linear programming [??], network flows [??], genetic algorithms [??], simulated annealing [??], and artificial intelligence [??].

@@

[...] Many of these approaches were designed to satisfy the requirements of a specific hospital department which causes a large number of variables and constraints to be incorporated into the problem formulation. While these department-specific approaches allow end-users to find precise schedules that satisfy the needs of the department and the preferences of the nurses and clinicians in that department, they are difficult to readily adapt to other departments in the same hospital or other hospitals. % explain why hard to adapt?... what makes their generalizablility/adaptability limited?
Moreover, the large number of variables and constraints also leads to computational complexity issues [ref], especially when using exact methods for finding the solution. In this paper, we tackle a version of the nurse scheduling problem arising from a case study of one clinical division, providing two different services simultaneously (general infectious disease (ID) consults; and HIV consults service) at St. Michael's Hospital in Toronto, Canada. Our goal is to (1) present a simple formulation for the problem developed and tested at the hospital after switching from a manual approach to scheduling; and (2) analyze the performance of integer linear programming in solving difficult instances of the problem and compare the results with those of the manual approach; and (3) describe the adaptability of the formulation as a basic framework for solving similar problems in other departments. \\ %make #3 an objective
We begin by describing the problem, then...% list the main contents/elements of the paper.

[...]

%Automated methods for scheduling in hospital settings have been studied extensively, most notably in the context of the Nurse Scheduling Problem (NSP) in Operations Research. In an NSP, the goal is to assign nurses to shifts, while satisfying various constraints pertaining to both hospital regulations and nurse preferences. The constraints can be classified as soft or hard constraints, based on the importance of satisfying them. Typically, it is necessary for a solution to satisfy all of the hard constraints, while the soft constraints are considered optional. The NSP, in its general case, is known to be NP-complete, meaning that although it is possible to efficiently check whether a given schedule satisfies all the constraints using an algorithm, it is not yet known whether an efficient algorithm for finding a satisfying schedule exists, and it is in fact equivalent to the hardest computational problems in Computer Science, Operations Research, and \textit{etc...} [??]. In order to tackle NP-complete problems such as NSP in real-life, algorithms either try to use various heuristics to approximate a potential solution rather than finding an exact solution or they focus on relatively small problem sizes, ensuring that the runtime of the algorithm is still practical. \textit{Some examples of heuristic approaches...} . \textit{Some examples of exact approaches (LP, etc.)} . A comprehensive review of various approaches is presented by Burke et al. [??]. \\

%In this paper we present the application of two methods, Network Flows and Integer Linear Programming (ILP), to a variation of the Nurse Scheduling Problem in order to generate a yearly schedule for clinicians working simultaneously at two different divisions of St. Michael's Hospital in Toronto, ON. The paper is organized as follows. Section [??] contains a detailed description of the scenario and a mathematical formalization of the problem. Section [??] describes first the Network Flow approach followed by the ILP approach to tackle the problem. Section [??] presents the results of both approaches when applied to the clinician data from St. Michael's Hospital. Section [??] presents results from simulations with a variety of problem sizes. We conclude the paper with a discussion of the real-world and simulation results in Section [??].

>>>>

Automated methods to optimize schedules have been studied and applied in many industries, including transportation [??], manufacturing [??], [...]. Of special interest to a clinician on-call scheduling problem are the approaches to schedule nurses, who often work in shifts. In the nurse scheduling problem, the goal is to find an optimal assignment of nurses to shifts that satisfies all of the hard constraints, such as hospital regulations, and as many soft constraints as possible, which may include nurse preferences. A wide variety of approaches, including exact and heuristic approaches, have been used to solve the nurse scheduling problem: integer linear programming [??], network flows [??], genetic algorithms [??], simulated annealing [??], and artificial intelligence [??]. \\

@@

Many of these approaches were designed to satisfy the requirements of a specific hospital department which causes a large number of variables and constraints to be incorporated into the problem formulation. While these department-specific approaches allow end-users to find precise schedules that satisfy the needs of the department and the preferences of the nurses and clinicians in that department, they are difficult to readily adapt to other departments in the same hospital or other hospitals. % explain why hard to adapt?... what makes their generalizablility/adaptability limited?
Moreover, the large number of variables and constraints also leads to computational complexity issues [ref], especially when using exact methods for finding the solution. In this paper, we tackle a version of the nurse scheduling problem arising from a case study of one clinical division, providing two different services simultaneously (general infectious disease (ID) consults; and HIV consults service) at St. Michael's Hospital in Toronto, Canada. Our goal is to (1) present a simple integer linear programming formulations for the scheduling problem as developed for the hospital, and describe the adaptability of the formulation to solving similar problems in other departments; (2) compare the performance of the ILP scheduler to the results of the manual approach; and (3) analyze the robustness of the algorithm in difficult instances of the problem. \\
% present a simple formulation for the problem developed and tested at the hospital after switching from a manual approach to scheduling; and (2) analyze the performance of integer linear programming in solving difficult instances of the problem and compare the results with those of the manual approach; and (3) describe the adaptability of the formulation as a basic framework for solving similar problems in other departments. \\ %make #3 an objective

We begin by describing the problem in Section \ref{sec:problem}, and presenting our ILP formulation in Section \ref{sec:methods}. Next, we compare the results of our formulation to manually-created schedules, and evaluate the performance of the algorithm on simulated data in Section \ref{sec:results}. Finally, we discuss and interpret the results in Section \ref{sec:discussion}. % list the main contents/elements of the paper.
